---
title: "Lab 8 -- Linear Regression in R"
author: "Joanna Tang"
date: "11/21/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load packages and read in data

library(tidyverse)
salmon <- read.csv("salmon_imports.csv")

```

First, let's look at our data.

```{r}

# Exploratory scatterplot
salmon_scatter <- ggplot(salmon, aes(x = year, y = salmon_fresh)) +
  geom_point()
salmon_scatter

# Overall, data looks relatively linear...there are just a couple outliers/weird dips
```

Now, let's fit a linear regression to our data.

```{r}

# Linear regression model
salmon_model <- lm(salmon_fresh ~ year, data = salmon)
salmon_model
# Imports(tons) = -14982940 + 7550(Year)
# This equation tells us that, for every year, salmon imports are increasing by 7550 tons.  This equation also extrapolates that, for "year 0", salmon imports were -14982940 tons -- this info is useless to us.  But that's the problem with models -- all are wrong, some are useful.

# Plot diagnostic plots of our model
par(mfrow = c(2,2))
plot(salmon_model)
# Residuals appear normally distributed
# Looks relatively homoscedastic (if we ignore some outliers)

# Let's look at a summary of our model
summary(salmon_model)
# p<.001, so we can reject the null hypothesis of no correlation between year salmon imports --> conclude that year significantly predicts salmon imports
# R^2 = .95, which means that 95% of salmon imports are predicted by year (this means that 5% of the variance in salmon imoprts is NOT explained by year)

```

Now let's create a final graph of the linear regression model with the original data, model, and 95% confidence interval for predictions.

```{r}
salmon_final_graph <- ggplot(salmon, aes(x = year, y = salmon_fresh)) +
  geom_point() +
  geom_smooth(method = lm, se = TRUE, size = .5, color = "grey20") +
  theme_bw() +
  scale_x_continuous(expand = c(0,0), limits = c(1990,2017)) +
  labs(x = "Year", y = "U.S. Annual Fresh Salmon Imports\n(1,000 pounds)")
salmon_final_graph

```

Now let's use our model to make predictions for US fresh salmon imports in year 2026 and 2028.  We'll do this by creating a data fram containing the "new values" that we want to make predictions for, and then we'll feed that into the predict() function.

```{r}

# Create new data frame with values we want to make predictions for
new_years <- data.frame(year = c(2022, 2024, 2025, 2028))
new_years

# Use predict() function
future_predict <- predict(salmon_model, newdata = new_years, interval = "confidence")
future_predict

predictions <- data.frame(new_years, future_predict)
predictions
```

Now let's find Pearson's r (correlation)

```{r}

salmon_r <- cor.test(salmon$year, salmon$salmon_fresh)
salmon_r

# Pearson's r = .974, p<.001 --> strong positive correlation between year and salmon imports
```

Year is a significant predictor of US fresh salmon imports (Imports(tons) = -14982940 + 7550(Year), *p* <0.001, Pearson's *r* = `r salmon_r$statistic`) and the overall model fit is *R^2^* = 0.95.